---
title: "Wild Catch or Fish Farm? How Australia's Seafood Industry Is Changing"
subtitle: "Assignment 4 ETC5512"
author: "Xuanning Zhang"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

# Data Details

## Main Question

How has the balance between wild-caught fisheries and aquaculture production in Australia changed over time?

## Sub-problem

1.  Which states are most responsible for the growth in aquaculture production?

2.  Has Australia undergone a structural shift from capture fisheries toward aquaculture at the national level since 2000?

### Reason

-   Significance: Australia is a major fishing nation, the research is related to food sustainability and marine environmental policy.

-   Personal: Interested to environmental policy and fishery.

## Data description

The datasets use for this research is to analyze the structural shifts between wild-caught fisheries and aquaculture production in Australia. For more details, including licence information and definitions for the processed data used in analysis, refer to the `README` file and `data-dictionary`.

### Dataset 1

**Source**

[Australian Bureau of Agricultural and Resource Economics and Sciences (ABARES), Department of Agriculture, Fisheries and Forestry.](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics)

**Content:** Annual production statistics reporting volume and value in AUD for wild-caught fisheries and aquaculture in national level and by state. The time range is from 1998-99 to 2023-24.

**Data Type:** Census

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0)

**Limitation:** State-level table may apply slightly different species groupings through years which will make long-run comparisons less reliable.

**Ethics & Privacy:** No individual or personal identifiable information included in this dataset, and this dataset is publicly available with no privacy concerns.

### Dataset 2

**Source:** [Australian Bureau of Statistics (ABS), Australian Statistical Geography Standard (ASGS).](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs/edition-3-july-2021-june-2026/access-and-downloads/digital-boundary-files)

**Content:** Digital boundary files for all Australian states and territories in 2021 edition.

**Data Type:** GIS data.

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0).

**Limitation:** Boundaries are based on 2021 edition, minor administrative boundary change these years are not reflected.

**Ethics & Privacy:** No privacy concerns, geographic boundary data only.

## Data Download and Processing

```{r, message=FALSE}
#load the required packages
library(tidyverse)
library(readxl)
library(sf)
library(here)
```

#### Data 1

The ABARES statistics file must download manually from ABARES portal:

1.  Visit [ABARES](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics).

2.  Download the statistical table (data to 2024-25) with the format of Excel.

3.  Save the data to **`data/raw/AFAS_Data_2024-25_v1.0.0`**

```{r}
excel_sheets(here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"))
```

**Processing national data**

```{r, warning=FALSE}
#read the national data in table 3
#skip top 7 rows
raw_t3 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 3",
  skip  = 7
)

#tidy out the commodity of wild-caught with the unit of per kilo AUD.
wild_national_val <- raw_t3 |> 
  slice(35) |> 
  select(-1, -2, -3) |> #first three colunm are useless
  pivot_longer(everything(),  #transfer to long format, easy to 
               names_to = "year",  #read
               values_to = "wildcatch_val",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_val)) #drop the na value in 2024-25

#read the quantity of wild-caught in the unit of tone
wild_national_qty <- raw_t3 |>
  slice(70) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_vol))


#read the aquaculture data in table 4
raw_t4 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 4",
  skip  = 7
)
#same way to calculate the commodity of aquaculture fishery
aqua_national_val <- raw_t4 |>
  slice(24) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |> filter(!is.na(aqua_val))
#calculate the quantity of aquacultre fishery
aqua_national_qty <- raw_t4 |>
  slice(48) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |> 
  filter(!is.na(aqua_vol))

#join all data together and save
national_clean <- wild_national_val |>
  left_join(aqua_national_val,  by = "year") |>
  left_join(wild_national_qty,  by = "year") |>
  left_join(aqua_national_qty,  by = "year") |>
  mutate(
    total_val           = wildcatch_val + aqua_val,
    aqua_share_val      = aqua_val / total_val,
    wildcatch_share_val = wildcatch_val / total_val,
    total_vol           = wildcatch_vol + aqua_vol,
    aqua_share_vol      = aqua_vol / total_vol,
    wildcatch_share_vol = wildcatch_vol/ total_vol
  )

#save the clean data
write_csv(national_clean, here("data", "processed", "national_production_clean.csv"))
```

**Processing data in state**

```{r, warning=FALSE}
#NSW
raw_t5 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 5", skip = 7
)
#value
nsw_wild_val <- raw_t5 |>
  slice(37) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "NSW")

nsw_auqa_val <- raw_t5 |>
  slice(49) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "NSW")

#quantity
nsw_wild_vol <- raw_t5 |>
  slice(87) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "NSW")

nsw_auqa_vol <- raw_t5 |>
  slice(99) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "NSW")

#VIC
raw_t6 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 6", skip = 7
)
#value
vic_wild_val <- raw_t6 |>
  slice(27) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "VIC")

vic_auqa_val <- raw_t6 |>
  slice(38) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "VIC")

#quantity
vic_wild_vol <- raw_t6 |>
  slice(66) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "VIC")

vic_auqa_vol <- raw_t6 |>
  slice(77) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "VIC")
#QLD
raw_t7 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 7", skip = 7
)
#value
qld_wild_val <- raw_t7 |>
  slice(38) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "QLD")

qld_auqa_val <- raw_t7 |>
  slice(50) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "QLD")

#quantity
qld_wild_vol <- raw_t7 |>
  slice(89) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "QLD")

qld_auqa_vol <- raw_t7 |>
  slice(101) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "QLD")

#SA
raw_t8 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 8", skip = 7
)
#value
sa_wild_val <- raw_t8 |>
  slice(29) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "SA")

sa_auqa_val <- raw_t8 |>
  slice(40) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "SA")

#quantity
sa_wild_vol <- raw_t8 |>
  slice(70) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "SA")

sa_auqa_vol <- raw_t8 |>
  slice(81) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "SA")

#WA
raw_t9 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 9", skip = 7
)
#value
wa_wild_val <- raw_t9 |>
  slice(35) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "WA")

wa_auqa_val <- raw_t9 |>
  slice(45) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "WA")

#quantity
wa_wild_vol <- raw_t9 |>
  slice(81) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "WA")

wa_auqa_vol <- raw_t9 |>
  slice(91) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "WA")

#TAS
raw_t10 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 10", skip = 7
)
#value
tas_wild_val <- raw_t10 |>
  slice(28) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "TAS")

tas_auqa_val <- raw_t10 |>
  slice(35) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "TAS")

#quantity
tas_wild_vol <- raw_t10 |>
  slice(64) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "TAS")

tas_auqa_vol <- raw_t10 |>
  slice(71) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "TAS")

#NT
raw_t11 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 11", skip = 7
)
#value
nt_wild_val <- raw_t11 |>
  slice(25) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "NT")

nt_auqa_val <- raw_t11 |>
  slice(30) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "NT")

#quantity
nt_wild_vol <- raw_t11 |>
  slice(56) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "NT")

nt_auqa_vol <- raw_t11 |>
  slice(61) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol),
         aqua_vol = replace_na(aqua_vol, 0)) |>
  filter(year != "2024–25p") |> 
  mutate(state = "NT")

#merge all states' value
all_wild_val <- bind_rows(
  nsw_wild_val,
  vic_wild_val, 
  qld_wild_val,
  sa_wild_val,
  wa_wild_val,
  tas_wild_val,
  nt_wild_val)

all_auqa_val <- bind_rows(
  nsw_auqa_val,
  vic_auqa_val,
  qld_auqa_val,
  sa_auqa_val,
  wa_auqa_val,
  tas_auqa_val,
  nt_auqa_val)
#merge all states' volumn
all_wild_vol <- bind_rows(
  nsw_wild_vol,
  vic_wild_vol,
  qld_wild_vol,
  sa_wild_vol,
  wa_wild_vol,
  tas_wild_vol,
  nt_wild_vol)

all_auqa_vol <- bind_rows(
  nsw_auqa_vol,
  vic_auqa_vol,
  qld_auqa_vol,
  sa_auqa_vol,
  wa_auqa_vol,
  tas_auqa_vol,
  nt_auqa_vol)

#join all data to form a clean dataset
state_clean <- all_wild_val |>
  left_join(all_auqa_val, by = c("year", "state")) |>
  left_join(all_wild_vol, by = c("year", "state")) |>
  left_join(all_auqa_vol, by = c("year", "state")) |>
  mutate(
    total_val      = wildcatch_val + aqua_val,
    aqua_share_val = aqua_val / total_val,
    total_vol      = wildcatch_vol + aqua_vol,
    aqua_share_vol = aqua_vol / total_vol
  )
#save dataset
write_csv(state_clean, here("data", "processed", "state_production_clean.csv"))
```

#### Data 2

The ABS boundary shapefile must be manually downloaded:

1.  Visit [ABS Digital Boundary Files](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs/edition-3-july-2021-june-2026/access-and-downloads/digital-boundary-files)

2.  Download **"State and Territory — 2021 — Shapefile"**.

3.  Unzip and save all files to `data/raw/STE_2021_AUST_SHP_GDA2020/`

**Data processing**

```{r}
#geomotry file does not need tidy, read only
states_sf <- st_read(
  here("data", "raw", "STE_2021_AUST_SHP_GDA2020",
       "STE_2021_AUST_GDA2020.shp")
)
```

<!-- Task 2 Tab: Writing your blogpost -->

# Blog post

## Wild Catch or Fish Farm? How Australia's Seafood Industry Is Changing

### What Are Wild-Caught Fisheries and Aquaculture?
Before delving into the data, it would be helpful to understand what we are comparing. Wild-caught fisheries refer to fish and seafood harvested directly from the ocean, rivers and lakes. Australia has one of the world's largest exclusive economic zoes which covering around 8.2 million square kilometers, and its whild-caught fisheries including a wide range of species from southern bluefin tuna in South Australian to rock lobsters off the coast of Western Australia (ABARES, 2025).

In contrast, aquaculture is the rearing of aquatic organisms in controlled environments like offshore net pens, inland ponds and onshore tanks. In Australia, Atlantic salmon farmed in the cold and deep waters of Tasmania dominates the sector in terms of production value, there also have alongside oysters, mussels and barramundi (ABARES, 2025).

Both industries are of significant economic importance. ABARES (2025) has pointed out that Australian fishing and aquaculture industries collectively produces over $ 3.8 billion output value during 2023-24 financial year which supporting fishing dependent communities across the country.

### Importance of Analytics on Balance of Wild-caught and Aquaculture
The impact of how a country obtains its seafood extends far beyond the dinner table. Aquaculture could provide a more stable food supply and reduce pressure on wild fish population in theory (Naylor et al., 2000).
However, large-scale salmon farming will bring environmental costs like nutrient pollution and biosecurity risks to wild populations (Buschmann et al., 2006). Also, the national shift towards aquaculture means communities which rely on wild-caught fish will gradually lose their economic.

### Data Analysis

```{r, echo=FALSE}
#| label: fig-area
national_long <- national_clean |>
  mutate(year_num = as.numeric(substr(year, 1, 4))) |>
  select(year_num, wildcatch_val, aqua_val) |>
  pivot_longer(cols = c(wildcatch_val, aqua_val),
               names_to  = "type",
               values_to = "value") |>
  mutate(
    type= ifelse(type== "wildcatch_val", "wild-caught", "Aquaculture"),
    value_billion= value/ 1000000
  )

ggplot(national_long, aes(x= year_num, y=value_billion, fill= type))+
  geom_area()+
  labs(
    title= "Australia's Fisheries Production Value (1998-2023)",
    x = "Year",
    y= "Production Value (AUD billion)"
  )+
  theme_minimal()
```

@fig-area show the total value of Australian fisheries production from 1998 to 2023 in financial years, the data has been separated into wild-caught and aquaculture. Wild capture was overwhelmingly dominant in late 1990s. However, the contribution of aquaculture grew steadily in las two decades while wild capture maintain largely unchanged. The value of aquaculture had surpassed that of wild caught which confirmed by ABARES (2025).

```{r, echo=FALSE}
#| label: fig-line
ggplot(national_clean, aes( x= as.numeric(substr(year, 1, 4)),
                            y= aqua_share_val))+
  geom_line(linewidth = 1)+
  geom_point(size=2)+
  scale_y_continuous(labels = scales::percent)+
  labs(title = "Aquaculture's Share of Total Fisheries Value in Australia",
       x= "Year",
       y= "Aquaculture Share (%)")+
  theme_minimal()
```

@fig-line has clearly point out the structural transformation. In 1998, there was only 28% of total fisheries value, wild-caught fisheries has dominated the industry by a wild margin. But in next 20 years, the aquaculture's share climed steady and reached around 57% of total by 2023 and firstly broke 50% in 2020 (ABARES, 2025).
This argument has been strengthened by @fig-vol. Even the production is measured by volume rather than value, aquaculture's share has confirmed that the transition is not the pricing effect driven by premium salmon prices, but reflect the structural growth in framed output.
In summary, both value and volume has pointed the same direction: Australia's seafood industry has shifted from wild-caught to aquaculture, and the trend shows no sign of reversing till now.
```{r, echo=FALSE}
#| label: fig-vol
ggplot(national_clean, aes(x = as.numeric(substr(year, 1, 4)),
                           y = aqua_share_vol)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title= "Aquaculture's Share of Total Fisheries Volume in Australia",
    x= "Year",
    y= "Aquaculture Share (%)"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
#| label: fig-map
#plot in map

#pick the last data
state_lastest <- state_clean |> 
  filter(year == max(year))
#create a table of states
state_tbl <- tibble(
  state = c("NSW", "VIC", "QLD", "SA", "WA", "TAS", "NT"),
  STE_NAME21 = c("New South Wales", "Victoria", "Queensland",
                 "South Australia", "Western Australia",
                 "Tasmania", "Northern Territory")
)
#join the table with full name and abbriviation of states
map_data <- state_lastest |> 
  left_join(state_tbl, by= "state") |> 
  right_join(states_sf, by= "STE_NAME21")

#draw the map
ggplot(map_data)+
  geom_sf(aes(fill=aqua_share_val, geometry= geometry))+
  labs(
    title= "Aquaculture Share of Total Fisheries Value by State (2023-24)",
    fill= "Aquaculture Share"
  )
```

```{r,echo=FALSE}
#| label: fig-dumbbell
#select the 2000-1 and 2023 data
state_dumbbell <- state_clean |>
  filter(year %in% c("2000–01", max(year))) |>
  select(state, year, aqua_share_val) |>
  mutate(period = ifelse(year== "2000–01", "Year 2000", "Year 2023-24"))
#transfer to wide format
state_wide <- state_dumbbell |>
  select(state, period, aqua_share_val) |>
  pivot_wider(names_from= period,
              values_from= aqua_share_val)
#create plot
ggplot(state_wide) +
  geom_segment(aes(x= `Year 2000`, xend = `Year 2023-24`,
                   y= state, yend = state),
               colour = "steelblue", linewidth = 1) +
  geom_point(aes(x= `Year 2000`, y = state), colour = "red", size = 4) +
  geom_point(aes(x= `Year 2023-24`, y = state), colour = "grey30", size = 4) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title= "Change in Aquaculture Share by State (2000–01 vs 2023–24)",
    x= "Aquaculture Share of Total Fisheries Value",
    caption= "Red = 2000–01, Grey = 2023–24"
  ) +
  theme_minimal()
```

1.  Analysis\
2.  Conclusions\
3.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

# Behind the Scenes

### What's in this section

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
