---
title: "What is your blogpost title? Add it here??"
subtitle: "Assignment 4 ETC5512"
author: "Who are you?? Add your name here"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

# Data Details

## Main Question

How has the balance between wild-caught fisheries and aquaculture production in Australia changed over time?

## Sub-problem

1.  Which species categories are most responsible for the growth in aquaculture production?

2.  Has Australia undergone a structural shift from capture fisheries toward aquaculture at the national level since 2000?

### Reason

-   Significance: Australian is a major fishing nation, the research could related with food sustainability and marine environmental policy.

-   Personal: Interested to environmental policy and fishery.

## Data description

The dataset use for this research is to analyze the structural shifts between wild-caught fisheries and aquaculture production in Australia. For more details, including licence information and definitions for the processed data used in analysis, refer to the `README` file and `data-dictionary`.

### Dataset 1

**Source**

[Australian Bureau if Agricultural and Resource Economics and Science (ABARES), Department of Agriculture, Fisheries and Forestry.](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics)

**Content:** Annual production statistics reporting volume and vale in AUD for wild-caught fisheries and aquaculture in national level and by state. The time range is from 1998-99 to 2023-2024.

**Data Type:** Census

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0)

**Limitation:** State-level table may apply slightly different species groupings through years which will make long-run comparisons less reliable.

**Ethics & Privacy:** No individual or personal identifiable information included in this dataset, and this dataset is publicly available with no privacy concerns.
:::

::: panel-tabset
### Dataset 2

**Source:** [Australian Bureau of Statistics (ABS), Australian Statistical Geography Standard (ASGS).](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs/edition-3-july-2021-june-2026/access-and-downloads/digital-boundary-files)

**Content:** Digital boundary files for all Australian states and territories in 2021 edition.

**Data Type:** GIS data.

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0).

**Limitation:** Boundaries are based on 2021 edition, minor administrative boundary change these years are not reflected.

**Ethics & Privacy:** No privacy concerns, geographic boundary data only.
:::

::: panel-tabset
## Data Download and Processing

```{r, message=FALSE}
#load the required packages
library(tidyverse)
library(readxl)
library(sf)
library(here)
```

**Dataset1**

The ABARES statistics file must download manually from ABARES portal:

1.  Visit [ABARES](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics).

2.  Download the statistical table (data to 2024-25) with the format of Excel.

3.  Save the data to **`data/raw/AFAS_Data_2024-25_v1.0.0`**

```{r}
excel_sheets(here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"))
```
:::

::: panel-tabset
```{r, warning=FALSE}
#read the national data in table 3
#skip top 7 rows
raw_t3 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 3",
  skip  = 7
)

#tidy out the commodity of wild-caught with the
#unit of per kilo AUD.
wild_national_val <- raw_t3 |> 
  slice(35) |> 
  select(-1, -2, -3) |> #first three colunm are useless
  pivot_longer(everything(),  #transfer to long format, easy to read
               names_to = "year",
               values_to = "wildcatch_val",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_val)) #drop the na value in 2024-25

#read the quantity of wild-caught in the unit of tone
wild_national_qty <- raw_t3 |>
  slice(70) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_vol))
```

<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

### What's in this section

Here is where you should write your blogpost! (Task 2)

### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)
2.  Motivation\
3.  Data\
4.  Analysis\
5.  Conclusions\
6.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

### What's in this section

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
