---
title: "Wild Catch or Fish Farm? How Australia's Seafood Industry Is Changing"
subtitle: "Assignment 4 ETC5512"
author: "Xuanning Zhang"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

# Data Details

## Main Question

How has the balance between wild-caught fisheries and aquaculture production in Australia changed over time?

## Sub-problem

1.  Which state are most responsible for the growth in aquaculture production?

2.  Has Australia undergone a structural shift from capture fisheries toward aquaculture at the national level since 2000?

### Reason

-   Significance: Australia is a major fishing nation, the research is related to food sustainability and marine environmental policy.

-   Personal: Interested to environmental policy and fishery.

## Data description

The dataset use for this research is to analyze the structural shifts between wild-caught fisheries and aquaculture production in Australia. For more details, including licence information and definitions for the processed data used in analysis, refer to the `README` file and `data-dictionary`.

### Dataset 1

**Source**

[Australian Bureau if Agricultural and Resource Economics and Sciences (ABARES), Department of Agriculture, Fisheries and Forestry.](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics)

**Content:** Annual production statistics reporting volume and value in AUD for wild-caught fisheries and aquaculture in national level and by state. The time range is from 1998-99 to 2023-24.

**Data Type:** Census

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0)

**Limitation:** State-level table may apply slightly different species groupings through years which will make long-run comparisons less reliable.

**Ethics & Privacy:** No individual or personal identifiable information included in this dataset, and this dataset is publicly available with no privacy concerns.

### Dataset 2

**Source:** [Australian Bureau of Statistics (ABS), Australian Statistical Geography Standard (ASGS).](https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs/edition-3-july-2021-june-2026/access-and-downloads/digital-boundary-files)

**Content:** Digital boundary files for all Australian states and territories in 2021 edition.

**Data Type:** GIS data.

**Licence:** Creative Commons Attribution 4.0 International (CC BY 4.0).

**Limitation:** Boundaries are based on 2021 edition, minor administrative boundary change these years are not reflected.

**Ethics & Privacy:** No privacy concerns, geographic boundary data only.

## Data Download and Processing

```{r, message=FALSE}
#load the required packages
library(tidyverse)
library(readxl)
library(sf)
library(here)
```

**Dataset1**

The ABARES statistics file must download manually from ABARES portal:

1.  Visit [ABARES](https://www.agriculture.gov.au/abares/research-topics/fisheries/fisheries-and-aquaculture-statistics).

2.  Download the statistical table (data to 2024-25) with the format of Excel.

3.  Save the data to **`data/raw/AFAS_Data_2024-25_v1.0.0`**

```{r}
excel_sheets(here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"))
```

**Processing national data**

```{r, warning=FALSE}
#read the national data in table 3
#skip top 7 rows
raw_t3 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 3",
  skip  = 7
)

#tidy out the commodity of wild-caught with the unit of per kilo AUD.
wild_national_val <- raw_t3 |> 
  slice(35) |> 
  select(-1, -2, -3) |> #first three colunm are useless
  pivot_longer(everything(),  #transfer to long format, easy to 
               names_to = "year",  #read
               values_to = "wildcatch_val",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_val)) #drop the na value in 2024-25

#read the quantity of wild-caught in the unit of tone
wild_national_qty <- raw_t3 |>
  slice(70) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol",
               values_transform = as.numeric) |> 
  filter(!is.na(wildcatch_vol))


#read the aquaculture data in table 4
raw_t4 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 4",
  skip  = 7
)
#same way to calculate the commodity of aquaculture fishery
aqua_national_val <- raw_t4 |>
  slice(24) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |> filter(!is.na(aqua_val))
#calculate the quantity of aquacultre fishery
aqua_national_qty <- raw_t4 |>
  slice(48) |>
  select(-1, -2, -3) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |> 
  filter(!is.na(aqua_vol))

#join all data together and save
national_clean <- wild_national_val |>
  left_join(aqua_national_val,  by = "year") |>
  left_join(wild_national_qty,  by = "year") |>
  left_join(aqua_national_qty,  by = "year") |>
  mutate(
    total_val           = wildcatch_val + aqua_val,
    aqua_share_val      = aqua_val / total_val,
    wildcatch_share_val = wildcatch_val / total_val,
    total_vol           = wildcatch_vol + aqua_vol,
    aqua_share_vol      = aqua_vol / total_vol,
    wildcatch_share_vol = wildcatch_vol/ total_vol
  )

#save the clean data
write_csv(national_clean, here("data", "processed", "national_production_clean.csv"))
```

**Processing data in state**

```{r, warning=FALSE}
#NSW
raw_t5 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 5", skip = 7
)
#value
nsw_wild_val <- raw_t5 |>
  slice(37) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "NSW")

nsw_auqa_val <- raw_t5 |>
  slice(49) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "NSW")

#quantity
nsw_wild_vol <- raw_t5 |>
  slice(87) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "NSW")

nsw_auqa_vol <- raw_t5 |>
  slice(99) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "NSW")

#VIC
raw_t6 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 6", skip = 7
)
#value
vic_wild_val <- raw_t6 |>
  slice(27) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "VIC")

vic_auqa_val <- raw_t6 |>
  slice(38) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "VIC")

#quantity
vic_wild_vol <- raw_t6 |>
  slice(66) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "VIC")

vic_auqa_vol <- raw_t6 |>
  slice(77) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "VIC")
#QLD
raw_t7 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 7", skip = 7
)
#value
qld_wild_val <- raw_t7 |>
  slice(38) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "QLD")

qld_auqa_val <- raw_t7 |>
  slice(50) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "QLD")

#quantity
qld_wild_vol <- raw_t7 |>
  slice(89) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "QLD")

qld_auqa_vol <- raw_t7 |>
  slice(101) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "QLD")

#SA
raw_t8 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 8", skip = 7
)
#value
sa_wild_val <- raw_t8 |>
  slice(29) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "SA")

sa_auqa_val <- raw_t8 |>
  slice(40) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "SA")

#quantity
sa_wild_vol <- raw_t8 |>
  slice(70) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "SA")

sa_auqa_vol <- raw_t8 |>
  slice(81) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "SA")

#WA
raw_t9 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 9", skip = 7
)
#value
wa_wild_val <- raw_t9 |>
  slice(35) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "WA")

wa_auqa_val <- raw_t9 |>
  slice(45) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "WA")

#quantity
wa_wild_vol <- raw_t9 |>
  slice(81) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "WA")

wa_auqa_vol <- raw_t9 |>
  slice(91) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "WA")

#TAS
raw_t10 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 10", skip = 7
)
#value
tas_wild_val <- raw_t10 |>
  slice(28) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "TAS")

tas_auqa_val <- raw_t10 |>
  slice(35) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "TAS")

#quantity
tas_wild_vol <- raw_t10 |>
  slice(64) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "TAS")

tas_auqa_vol <- raw_t10 |>
  slice(71) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol)) |>
  filter(!is.na(aqua_vol)) |>
  mutate(state = "TAS")

#NT
raw_t11 <- read_xlsx(
  here("data", "raw", "AFAS_Data_2024-25_v1.0.0.xlsx"),
  sheet = "Table 11", skip = 7
)
#value
nt_wild_val <- raw_t11 |>
  slice(25) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_val") |>
  mutate(wildcatch_val = as.numeric(wildcatch_val)) |>
  filter(!is.na(wildcatch_val)) |>
  mutate(state = "NT")

nt_auqa_val <- raw_t11 |>
  slice(30) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_val") |>
  mutate(aqua_val = as.numeric(aqua_val)) |>
  filter(!is.na(aqua_val)) |>
  mutate(state = "NT")

#quantity
nt_wild_vol <- raw_t11 |>
  slice(56) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |> 
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "wildcatch_vol") |>
  mutate(wildcatch_vol = as.numeric(wildcatch_vol)) |>
  filter(!is.na(wildcatch_vol)) |>
  mutate(state = "NT")

nt_auqa_vol <- raw_t11 |>
  slice(61) |>
  select(-1, -2, -3) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(everything(),
               names_to  = "year",
               values_to = "aqua_vol") |>
  mutate(aqua_vol = as.numeric(aqua_vol),
         aqua_vol = replace_na(aqua_vol, 0)) |>
  filter(year != "2024–25p") |> 
  mutate(state = "NT")

#merge all states' value
all_wild_val <- bind_rows(
  nsw_wild_val,
  vic_wild_val, 
  qld_wild_val,
  sa_wild_val,
  wa_wild_val,
  tas_wild_val,
  nt_wild_val)

all_auqa_val <- bind_rows(
  nsw_auqa_val,
  vic_auqa_val,
  qld_auqa_val,
  sa_auqa_val,
  wa_auqa_val,
  tas_auqa_val,
  nt_auqa_val)
#merge all states' volumn
all_wild_vol <- bind_rows(
  nsw_wild_vol,
  vic_wild_vol,
  qld_wild_vol,
  sa_wild_vol,
  wa_wild_vol,
  tas_wild_vol,
  nt_wild_vol)

all_auqa_vol <- bind_rows(
  nsw_auqa_vol,
  vic_auqa_vol,
  qld_auqa_vol,
  sa_auqa_vol,
  wa_auqa_vol,
  tas_auqa_vol,
  nt_auqa_vol)

#join all data to form a clean dataset
state_clean <- all_wild_val |>
  left_join(all_auqa_val, by = c("year", "state")) |>
  left_join(all_wild_vol, by = c("year", "state")) |>
  left_join(all_auqa_vol, by = c("year", "state")) |>
  mutate(
    total_val      = wildcatch_val + aqua_val,
    aqua_share_val = aqua_val / total_val,
    total_vol      = wildcatch_vol + aqua_vol,
    aqua_share_vol = aqua_vol / total_vol
  )
#save dataset
write_csv(state_clean, here("data", "processed", "state_production_clean.csv"))
```

<!-- Task 2 Tab: Writing your blogpost -->

# Blog post

### What's in this section

Here is where you should write your blogpost! (Task 2)

### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)
2.  Motivation\
3.  Data\
4.  Analysis\
5.  Conclusions\
6.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

# Behind the Scenes

### What's in this section

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
